\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-02-23 forwards-and-backwards.tex
alison edits feb 21

If one replaces the logistic growth term of the classical
Fisher-KPP equation with a nonlinearity that reflects cooperative
behaviour in the population, such as
\begin{equation}
	\label{Birzu nonlinearity}
	wF(w)=w(1-w)(Cw-1),
\end{equation}
then, for sufficiently large $C$ (strong cooperation),
the nature of the deterministic
wave changes from ``pulled'' to ``pushed'', \cite{birzu/hallatschek/korolev:2017}.
In that setting, the genealogies will be quite different
from the Fisher-KPP case. \comment{Could refer to more recent Birzu et al paper here.}
For example, \cite{etheridge/penington:2022}
show that for a discrete space model corresponding to this 
nonlinearity with $C>2$, after suitable scaling, the genealogy of a
sample converges not to a Bolthausen-Sznitman coalescent, but to
a Kingman coalescent. 
The reason, roughly, is that ancestral lineages
settle to a stationary distribution relative to the position of the 
wavefront which puts very little weight close to the `tip' of the wave, so
that when ancestral lineages meet
it is typically at a location in which population density is
high, where no single ancestor produces a disproportionately large number of 
descendants in a short space of time.

@@

For each $N$ and $\theta$, the rescaled process
$\big(\eta^N_t\big)_{t\geq 0} := \big(X(\theta t)/N\big)_{t\geq 0}$, 
takes values in the space of c\`adl\`ag paths in
$\measures$ (the space of finite measures on $\IR^d$
endowed with the weak topology).

@@

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists. From the 
expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limit points will be deterministic or stochastic processes 
is determined by whether $\alpha$ is zero or nonzero.
%c.f.~Section~\ref{sec:heuristics}.

%In order to specify the assumptions that we shall need on the parameters of our model
%if we are to pass to a scaling limit, 
It is convenient to record some notation for the 
generator of the diffusion 
limit of a random walk with jump distribution determined by $q_\theta(x,dy)$.

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although sometimes we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        %\text{TODO} .
    \end{align}
\end{definition}

@@

To keep expressions manageable, we shall also assume that 
\[
\mu_\theta(x) = r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N),
\]
that is, this expression is non-negative so that there is no need to take the maximum
with zero in~(\ref{eqn:mu_defn}).
\end{assumptions}

@@

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
\comment{Not sure next sentence strictly true - probably assuming a bit
more - certainly some regularity.}
Condition~\ref{control through r} is satisfied if 
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This will be satsified if $\rho_r$ decays 
no faster than
exponentially.
On the other hand, it might, for example, seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{old 6.1}, \comment{this is referring to the
old Lemma 6.1, now to go in Appendix,}
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

\paragraph{Scaling limits with nonlocal interactions}

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

Recall the following notion of solution to a PDE.
\begin{definition}[Weak solutions]
    \label{defn:weak_solutions}
    We say that $(\eta_t)_{t \ge 0}$, with $\eta_t \in \measures$,
    is a \emph{weak solution} to the PDE
   \[ 
	\dot \varphi = r \DG^*(\gamma \varphi) + \varphi F
    \]
	(where $r$, $\gamma$ and $F$ can all be functions of $\varphi$)
    if, for all $f \in C_b^\infty(\IR^d)$,
    \[
        \frac{d}{dt} \langle f, \eta_t \rangle
        =
        \langle
	    \gamma \DG(rf) + f F,
            \eta_t
        \rangle .
    \]
\end{definition}
the notation $\varphi$ is meant to be suggestive of a density.
Because Theorem~\ref{thm:nonlocal_convergence} only tells us about 
weak convergence, in the case $\alpha=0$ we can only deduce that 
any limit point $\eta_t$ is a weak solution to the nonlocal PDE
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi_t(x)
        =
        r\left(x, \smooth{r} \varphi_t(x) \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \smooth{F} \varphi_t(x) \right)
        .
    \end{equation}

@@

We write 
$\gamma(x,\eta)C(x)=J(x,\eta)J(x,\eta)^T$, and 
$\beta(x,\eta)=\gamma(x,\eta)\big(\meanq(x)+2C(x)\nabla\log r(x,\eta)\big)$.
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
%that
%\[|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
%	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
%	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))\]
%for some $C>0$.  
%
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

\paragraph{Two-step convergence to PDE}

>>>>

If one replaces the logistic growth term of the classical
Fisher-KPP equation with a nonlinearity that reflects cooperative
behaviour in the population, such as
\begin{equation}
	\label{Birzu nonlinearity}
	wF(w)=w(1-w)(Cw-1),
\end{equation}
then, for sufficiently large $C$ (strong cooperation),
the nature of the deterministic
wave changes from ``pulled'' to ``pushed'', 
\cite{birzu/hallatschek/korolev:2017, birzu/hallatschek/korolev:2021}.
In that setting, the genealogies will be quite different
from the Fisher-KPP case. 
%\comment{Could refer to more recent Birzu et al paper here.}
For example, \cite{etheridge/penington:2022}
show that for a discrete space model corresponding to this 
nonlinearity with $C>2$, after suitable scaling, the genealogy of a
sample converges not to a Bolthausen-Sznitman coalescent, but to
a Kingman coalescent. 
The reason, roughly, is that ancestral lineages
settle to a stationary distribution relative to the position of the 
wavefront which puts very little weight close to the `tip' of the wave, so
that when ancestral lineages meet
it is typically at a location in which population density is
high, where no single ancestor produces a disproportionately large number of 
descendants in a short space of time.

@@

For each $N$ and $\theta$, the rescaled process
$\big(\eta^N_t\big)_{t\geq 0} := \big(X(\theta t)/N\big)_{t\geq 0}$, 
takes values in the space of c\`adl\`ag paths in
$\measures$ (the space of finite measures on $\IR^d$
endowed with the weak topology). In fact $\eta_t^N$ will be a purely atomic
measure comprised of atoms of mass $1/N$.

@@

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists. From the 
expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limit points will be deterministic or stochastic processes 
is determined by whether $\alpha$ is zero or nonzero.

It is convenient to record some notation for the 
generator of the diffusion 
limit of a random walk with jump distribution determined by $q_\theta(x,dy)$.

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        %\text{TODO} .
    \end{align}
\end{definition}

@@

To keep expressions manageable, we shall also assume that 
\[
\mu_\theta(x) = r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N),
\]
that is, this expression is non-negative so that there is no need to take the maximum
	with zero in~(\ref{eqn:mu_defn}). (This is anyway true for 
	sufficiently large $\theta$.)
\end{assumptions}

@@

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
Condition~\ref{control through r} is satisfied if 
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This will be satsified, for example, if $\rho_r$ decays 
exponentially.
On the other hand, it might, for example, seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{old 6.1}, \comment{this is referring to the
old Lemma 6.1, now to go in Appendix,}
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

\paragraph{Scaling limits with nonlocal interactions:}

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with $\eta^N$ comprised of atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

Recall the following notion of solution to a PDE.
\begin{definition}[Weak solutions]
    \label{defn:weak_solutions}
    We say that $(\eta_t)_{t \ge 0}$, with $\eta_t \in \measures$,
    is a \emph{weak solution} to the PDE
   \[ 
	\partial_t \varphi = r \DG^*(\gamma \varphi) + \varphi F
    \]
	(where $r$, $\gamma$ and $F$ can all be functions of $\varphi$)
    if, for all $f \in C_b^\infty(\IR^d)$,
    \[
        \frac{d}{dt} \langle f, \eta_t \rangle
        =
        \langle
	    \gamma \DG(rf) + f F,
            \eta_t
        \rangle .
    \]
\end{definition}
the notation $\varphi$ is meant to be suggestive of a density.
Because Theorem~\ref{thm:nonlocal_convergence} only tells us about 
weak convergence, in the case $\alpha=0$ we can only deduce that 
any limit point $\eta_t$ is a weak solution to the nonlocal PDE
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi_t(x)
        =
        r\left(x, \smooth{r} \varphi_t(x) \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \smooth{F} \varphi_t(x) \right)
        .
    \end{equation}

@@

We write 
$\gamma(x,\eta)C(x)=J(x,\eta)J(x,\eta)^T$, and 
$\beta(x,\eta)=\gamma(x,\eta)\big(\meanq(x)+2C(x)\nabla\log r(x,\eta)\big)$.
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

\paragraph{Two-step convergence to PDE:}

Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula. 
\begin{proposition}
    \label{prop:nonlocal_to_local}
Recall the notation $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$b(x)$ and $C(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty)>0$,
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}
\begin{remark}
Note that Theorem~\ref{thm:kurtzxiong} guarantees uniqueness of solutions
to equation~(\ref{nonlocalPDEv1}). 
\end{remark}
Our second example is a nonlocal version of a porous medium equation with
logistic growth. That is, 
we consider non-negative solutions to the equation
\begin{equation}
\label{mollified equation}
\partial_t \psi^\epsilon=
\Delta\left(\psi^\epsilon \, \rho^{\epsilon}_\gamma*\psi^\epsilon\right)
+\psi^\epsilon\left(1-\rho^{\epsilon}_\gamma*\psi^\epsilon\right).
\end{equation}
The case without the reaction term (and with $\IR^d$ replaced by a torus) is considered 
by~\cite{lions/mas-gallic:2001} who use
it as a basis for a particle method for numerical solution of the porous medium equation.
Of course this does not quite fit into our framework, since in the notation of our 
population models this would necessitate $\gamma(x,m)=\smooth{\epsilon}m$ which is
not bounded. However, this can be overcome by an additional layer of approximation
(c.f.~our numerical experiments of Section~\ref{beyond linear diffusion})
and we do not allow this to detain us here. Existence and uniqueness of solutions 
to~(\ref{mollified equation}) can be obtained using the approach
of~\cite{lions/mas-gallic:2001}.
